---
title: "Data Transformation Solutions"
output: github_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The following questions are based on concepts covered in [Chapter 5][chapter 5]
of R4DS, and answers to them lie in the `transactions` dataset of the
completejourney package. Load the tidyverse and completejourney packages to
start working on them.

```{r load-packages, warning = FALSE, message = FALSE}
library(tidyverse)
library(completejourney)
```

**Note**: If printing the completjourney package's datasets to the console is
throwing you a warning that 'along' is being autocompleted as 'along.with', you
might want to toggle `warnPartialMatchArgs` off. Jenny Bryan writes that she has
been doing this a lot of late [here][jenny on github]. When the next version of
R is released, the bug causing this error will be fixed. For now, you can toggle
`warnPartialMatchArgs` off with:

```{r}
options(warnPartialMatchArgs = FALSE)
```

---

**Question 1**: Change the discount variables (i.e., `retail_disc`, 
`coupon_disc`, `coupon_match_disc`) from negative to positive. 

**Hint:** Use the `abs()` function within [`mutate()`][mutate]. 

We designed this question to strengthen your ability to use the dplyr verb
[`mutate()`][mutate] for the purposes of overwriting existing variables.

```{r include = TRUE}
transactions <- transactions %>% 
  mutate(
    retail_disc       = abs(retail_disc),
    coupon_disc       = abs(coupon_disc),
    coupon_match_disc = abs(coupon_match_disc)
  )
```

---

**Question 2**: Create three new variables named `regular_price`, 
`loyalty_price`, and `coupon_price` according to the following logic:

```{r eval = FALSE}
regular_price = (sales_value + retail_disc + coupon_match_disc) / quantity
loyalty_price = (sales_value + coupon_match_disc) / quantity
coupon_price  = (sales_value - coupon_disc) / quantity
```

We designed this question to strengthen your ability to use the dplyr verb
[`mutate()`][mutate] to create new variables. It should also help you develop a
better understanding of the discount variables in `transactions`.

```{r include = TRUE}
(transactions <- transactions %>% 
  mutate(
    regular_price = (sales_value + retail_disc + coupon_match_disc) / 
                     quantity,
    loyalty_price = (sales_value + coupon_match_disc) / 
                     quantity,
    coupon_price  = (sales_value - coupon_disc) / 
                     quantity
  )
)
```

---

**Question 3**: The `transactions` dataset includes 68,509 unique product IDs.
How many of these products (not transactions!) had a regular price of one dollar
or less? What does this count equal for loyalty price and coupon price?

**Hint:** After filtering, select the `product_id` column, then count the number
of unique products using the `n_distinct()` function.

We designed this question to strengthen your ability to use the dplyr verbs
[`filter()`][filter] and [`select()`][select].

```{r include = TRUE}
transactions %>% 
  filter(regular_price <= 1) %>% 
  select(product_id) %>% 
  n_distinct()

# 8,698 products had a regular price less than or equal to $1.00.

transactions %>% 
  filter(loyalty_price <= 1) %>% 
  select(product_id) %>% 
  n_distinct()

# This count for loyalty price is 14,043 products.

transactions %>% 
  filter(coupon_price <= 1) %>% 
  select(product_id) %>% 
  n_distinct()

# For coupon price, it's higher still at 15,676. 

# These numbers indicate that over 5,300 products were discounted to a dollar or
# less with loyalty-card promotions or coupons.
```

---

**Question 4**: What proportion of baskets are over $10 in sales value? 

**Hint:** You need to use [`group_by()`][summarize], [`summarize()`][summarize],
and [`ungroup()`][ungroup]. As a last step, you can calculate the proportion by
taking the mean of `TRUE/FALSE` values, using `mean(basket_value > 10)` to get
the proportion over $10.

We designed this question to strengthen your ability to use the dplyr verbs
[`group_by()`][summarize], [`summarize()`][summarize] and
[`ungroup()`][ungroup].

```{r include = TRUE}
transactions %>%
  group_by(basket_id) %>%
  summarize(basket_value = sum(sales_value)) %>%
  ungroup() %>%
  summarize(proportion_over_10 = round(mean(basket_value > 10) * 100, 0))

# Approximately 66% of customers' baskets bring the retailer over $10 in value.
```

---

**Question 5**: Which store with over $10K in total `sales_value` discounts its 
products the most for loyal customers? 

**Hint:** You can calculate loyalty discount as a percentage of regular price 
using the following logic:

```{r eval = FALSE}
pct_loyalty_disc = 1 - (loyalty_price / regular_price)
```

We designed this question to strengthen your ability to use the dplyr verbs 
[`filter()`][filter], [`mutate()`][mutate], [`group_by()`][summarize], 
[`summarize()`][summarize], and [`arrange()`][arrange] (i.e., 
almost everything covered in [Chapter 5][chapter 5] of R4DS).

```{r include = TRUE}
transactions %>%
  filter(
    is.finite(regular_price), 
    is.finite(loyalty_price), 
    regular_price > 0
  ) %>%
  mutate(
    pct_loyalty_disc     = 1 - (loyalty_price / regular_price)
  ) %>%
  group_by(store_id) %>%
  summarize(
    total_sales_value    = sum(sales_value), 
    avg_pct_loyalty_disc = mean(pct_loyalty_disc)
  ) %>%
  filter(total_sales_value > 10000) %>%
  arrange(desc(avg_pct_loyalty_disc))

# store_id == 341 has, on average, a discount of 18.7% off regular prices for
# members of the retailer's loyalty program. The code below returns all stores
# in rank order to give a sense of how (dis)similar the discounting is between
# stores.
```

[arrange]: http://r4ds.had.co.nz/transform.html#arrange-rows-with-arrange
[chapter 5]: http://r4ds.had.co.nz/transform.html
[filter]: http://r4ds.had.co.nz/transform.html#filter-rows-with-filter
[jenny on github]: https://github.com/tidyverse/tidyr/issues/519#issuecomment-439148810
[mutate]: http://r4ds.had.co.nz/transform.html#add-new-variables-with-mutate
[r4ds]: http://r4ds.had.co.nz/
[select]: http://r4ds.had.co.nz/transform.html#select-columns-with-select
[summarize]: http://r4ds.had.co.nz/transform.html#grouped-summaries-with-summarise
[ungroup]: http://r4ds.had.co.nz/transform.html#ungrouping