---
title: "Exploratory Data Analysis (EDA) Exercises"
author: "Zoe Zheng"
date: April 10, 2018
output: github_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, eval = TRUE, message = TRUE)
pdf(NULL)
```

The following seven exercises are based on concepts covered in 
[Chapter 7 of R4DS][G&W 2017, 7]. They can be answered using datasets from the
`completejourney` package. Start by loading the `tidyverse` and 
`completejourney` packages.

```{r warning = FALSE, message = FALSE}
library(tidyverse)
library(completejourney)
```

After running the following code, use `my_transaction_data` to work on the
following exercises.

```{r echo = TRUE}
left_join(transaction_data, product) %>% 
   left_join(hh_demographic) %>% 
   filter(
     quantity != 0
   ) %>% 
   mutate(
     regular_price  = (sales_value + retail_disc + coupon_match_disc) /
                       quantity,
     loyalty_price  = (sales_value + coupon_match_disc) / 
                       quantity,
     coupon_price   = (sales_value - coupon_disc) / 
                       quantity,
     purchase_price = ifelse(coupon_disc > 0, coupon_price, 
                             ifelse(retail_disc > 0, loyalty_price,
                                                            regular_price))
  ) -> my_transaction_data
```

---

## Exercise 1
How many unique households exist in `my_transaction_data`, and how many of these
households in `my_transaction_data` have demographic data in `hh_demographic`?

 a. Use `distinct()` to create a tibble of unique `household_key` values.
 b. Use `nrow()` to count these households.
 c. Use `inner_join()` to match `my_transaction_data` with `hh_demographic`.
 d. Use `distinct()` and `nrow()` to count the rows that remain.

```{r question 1, eval=TRUE}
my_transaction_data %>% 
  distinct(household_key) %>% 
  nrow()
# There are 2500 unique households in my_transaction_data

inner_join(my_transaction_data, hh_demographic) %>% 
  distinct(household_key) %>% 
  nrow()
# There are 801 households in my_transaction_data that have demographic data.
```
 
---

## Exercise 2

Determine median weekly spend per individual using the following tibble (i.e.,
`exercise_2`).

```{r}
inner_join(my_transaction_data, hh_demographic) %>% 
   mutate(
     hh_size          = str_replace(household_size_desc, '5\\+', '5') %>% 
                         as.integer()
   ) %>% 
   group_by(household_key, week_no) %>% 
   summarize(
     total_spend      = sum(purchase_price, na.rm = TRUE),
     hh_size          = max(hh_size,        na.rm = TRUE)
   ) %>% 
   ungroup() %>%
  mutate(
    wkly_spend_per_ind = total_spend / hh_size
  ) -> exercise_2
```



```{r question 2}
exercise_2 %>% 
  summarize(
    med_wkly_spend_per_ind = median(wkly_spend_per_ind, na.rm = TRUE)
  )
```

The median weekly spend per individual is $19.425.

---

## Exercise 3
Building on Exercise 2, plot median spend per individual for the five household
sizes in `my_transaction_data`.

```{r question 3}
exercise_2 %>% 
  group_by(hh_size) %>% 
  summarize(
    med_wkly_spend_per_ind = median(wkly_spend_per_ind, na.rm = TRUE)
  ) %>% 
  ggplot(aes(x = hh_size, y = med_wkly_spend_per_ind)) +
  geom_col()
```

---

## Exercise 4
Are baskets with diapers in them more likely than average to have beer in them
too? Legend has it that placing these two product categories closer together can
increase beer sales [(Powers 2002)][Powers 2002]. Using the following starter
code, calculate [lift][Wikipedia, lift] for the "association rule" that diapers
in a basket (i.e., `sub_commodity_desc == 'BABY DIAPERS'`) imply beer is in the
basket (i.e., `sub_commodity_desc == 'BEERALEMALT LIQUORS'`). Is the association
between these products practically significant in `my_transaction_data`?

```{r }
inner_join(my_transaction_data, product) %>% 
  mutate(
    diapers = sub_commodity_desc == 'BABY DIAPERS', 
    beer    = sub_commodity_desc == 'BEERALEMALT LIQUORS'
  )
```

```{r question 4}
inner_join(my_transaction_data, product) %>% 
  mutate(
    diapers = sub_commodity_desc == 'BABY DIAPERS', 
    beer    = sub_commodity_desc == 'BEERALEMALT LIQUORS'
  ) %>%
  group_by(basket_id) %>%
  summarize(
    basket_has_diapers = max(diapers), 
    basket_has_beer    = max(beer)
  ) %>% 
  summarize(
    prop_both   = sum(basket_has_diapers * basket_has_beer == 1) / 
                  sum(basket_has_diapers == 1),
    prob_beer   = mean(basket_has_beer),
    diaper_lift = prop_both / prob_beer
  )
```

diaper_lift>1, meaning that the probability of buying both diapers and beer is 
higher than probability of just buying beer. So, there is a significant association
between these products.


---

## Exercise 5
Using a stacked bar chart that's partitioned by income level (i.e.,
`income_desc`), visualize the total amount of money customers spent on
national-brand products versus private-label products.

```{r question 5}
inner_join(my_transaction_data, hh_demographic) %>% 
  mutate(
    income_desc = factor(income_desc, 
                         levels = c('Under 15K',   '15-24K',   '25-34K', 
                                       '35-49K',   '50-74K',   '75-99K', 
                                     '100-124K', '125-149K', '150-174K', 
                                     '175-199K', '200-249K',    '250K+'),
                         ordered = TRUE)
  ) %>%
  group_by(income_desc, brand) %>%
  summarize(total_spend = sum(purchase_price)) %>% 
  ggplot() +
  geom_col(aes(x = income_desc, y = total_spend, fill = brand), 
           position = 'fill') + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

Overall, consumers across all income level spend more on national-brand products
than private-label products. 

[G&W 2017, 7]: http://r4ds.had.co.nz/exploratory-data-analysis.html
[Powers 2002]: https://www.theregister.co.uk/2006/08/15/beer_diapers/
[Wikipedia, lift]: https://en.wikipedia.org/wiki/Lift_(data_mining)