---
#FilipKawka
title: "Data Wrangling Solutions"
output: github_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  include = TRUE,
  eval    = TRUE,
  warning = FALSE, 
  message = FALSE
)
```

The following questions are based on concepts covered in [Chapter 12][chapter
12] and [Chapter 13][chapter 13] of R4DS, and answers to them lie in the
`coupons`, `coupon_redemptions`, `campaigns`, `transactions`, and `products`
datasets of the completejourney package. Load the tidyverse, completejourney,
and lubridate packages to start working on them.

```{r}
library(tidyverse)
library(completejourney)
library(lubridate) # see chapter 16 of r4ds
```

---

**Question 1**: What percent of households that received the retailer's weekly
mailer redeemed at least one coupon?

```{r}
left_join(
  campaigns          %>% count(household_id, name = "n_recipients"),
  coupon_redemptions %>% count(household_id, name = "n_redemptions"), 
  by = "household_id"
) %>% 
  summarize(redemption_rate = mean(!is.na(n_redemptions)))

# Of the 1,559 households that were mailed coupons as part of a campaign, 410
# (or 26.3%) redeemed a coupon.
```

---

**Question 2**: How many households received and did not redeem a coupon?   

```{r}
left_join(
  campaigns          %>% count(household_id, name = "n_recipients"),
  coupon_redemptions %>% count(household_id, name = "n_redemptions"), 
  by = "household_id"
) %>% 
  summarize(redemption_rate = sum(is.na(n_redemptions)))

# 1,149 households received and did not redeem a coupon.
```

---

**Question 3**: What percentage of coupons promoted in the retailer's weekly
mailer got redeemed at least once?

```{r}
left_join(
      coupons            %>% count(coupon_upc, name = "n_products", sort = TRUE),
      coupon_redemptions %>% count(coupon_upc, name = "n_redemptions"), 
      by = "coupon_upc") %>% 
  summarize(redemption_rate = sum(!is.na(n_redemptions)))

# Of the 981 coupons promoted in the retailer's weekly mailer, 491 (or 50%) of
# them were redeemed at least once.
```

---

**Question 4**: Considering the product categories that the 801 households in
the Complete Journey Study purchased most heavily, which five categories did
they start spending more on at the highest rate over the course of Q1? Only
consider product categories that the group spent $1,500 or more on in January,
and calculate spend growth as a percentage of category spend in January.

Here are some suggestions:  
 1. Join the `transactions` and `products` datasets.  
 2. Get month from `transaction_timestamp` and name this variable `month`.  
 3. Group by `product_category` and `month`.  
 4. Calculate total spend at the category-month level using `sales_value`.  
 5. Filter the data to include only the months of January and March.  
 6. Create a new variable called `spend_growth_pct`, making use of the
 [`lag()`][5.5.1] function.  
 7. Filter for categories that had category spend of \$1,500 or more in January.  
 8. Arrange the data in descending order according to `spend_growth_pct`.  

```{r}
transactions %>% 
  left_join(products, by = "product_id") %>%
  mutate(month = month(transaction_timestamp, label = TRUE)) %>%
  group_by(product_category, month) %>%
  summarize(spend_tot = sum(sales_value, na.rm = TRUE)) %>%
  filter(month == "Jan" | month == "Mar") %>%
  group_by(product_category) %>% 
  mutate(
    spend_growth_pct = (spend_tot - lag(spend_tot)) / lag(spend_tot) * 100,
    spend_jan = first(spend_tot)
  ) %>% 
  filter(spend_jan >= 1500) %>% 
  arrange(desc(spend_growth_pct)) %>% 
  select(product_category, spend_growth_pct) %>% 
  head(5)

# From January to March, the 2,469 households in the Complete Journey Study
# increased their spend on infant formula by 53.4 percent. They spent almost 30
# percent more on frozen seafood, just over 23 percent more on packaged candy,
# nearly 20 percent more on oral hygiene products, and 13.5 percent more on
# domestic wine.
```

[5.5.1]: https://r4ds.had.co.nz/transform.html#mutate-funs
[r4ds]: http://r4ds.had.co.nz/
[chapter 12]: http://r4ds.had.co.nz/tidy-data.html
[chapter 13]: http://r4ds.had.co.nz/relational-data.html