---
title: "Brandefy Private Label Case Solutions"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The following 7 questions are based on the Brandefy case study and can be answered 
using The Complete Journey data. The Complete Journey data is an ideal test bed for 
better understanding consumer behavior around national and private label brands 
as Meg Greenhalgh considers building a Brandefy app. There is a column entitled "brand" 
in the `product` dataset that shows for each product whether it is a "National" 
or "Private" brand. Start your analysis by loading the `tidyverse`, `scales`, 
`gridExtra`, and `completejourney` packages. 

```{r load-packages, warning=FALSE, message=FALSE}
# required packages
library(tidyverse)
library(gridExtra)
library(scales)
library(completejourney)
```

---

**Question 1**:  
Meg Greenhalgh is trying to gauge the overall opportunity for usage of an app that shoppers 
could use in-store to scan items and see product reviews of similar private label 
products. Meg Greenhalgh thinks that shoppers are confronted with roughly 10 opportunities 
per trip to decide between a national or private-label brand. Using the Complete Journey 
dataset count the number of private label items per basket and plot the distribution 
as a histogram. This will help to start to get a sense for how many times consumers 
purchase a private label item per trip. Hint: When calculating the private label 
items per basket ensure you include the baskets with zero private label items. 
You can do this by summarizing the baskets using: 
`summarize(private_item_cnt = sum(brand == 'Private'))`.

**Answer**:  
In the histogram it appears that the majority of baskets have a low number of private 
label items; however, it is a very long-tailed distribution. In order to better gauge 
a rough estimate from the plot we've provided a histogram ignoring baskets with 
greater than 30 private label items. The long tail is an interesting insight because 
Meg Greenhalgh may be able to sub-segment potential users of the app into casual 
buyers and die-hard private label fans that would purchase private label items in bulk.

```{r Q1, warning=FALSE, message=FALSE}
private_item_cnt <- transaction_data %>%
  inner_join(product, by='product_id') %>%
  group_by(basket_id) %>%
  summarize(private_item_cnt = sum(brand == 'Private'))

plot1 <- ggplot(private_item_cnt) + 
  geom_histogram(aes(private_item_cnt), binwidth=3) + 
  xlab("Private Label Items in Basket") + 
  ylab("Count") + 
  scale_y_continuous(labels=comma) + 
  ggtitle(label = "Distribution of\nPrivate Label Items\nper Basket", 
          subtitle = "Considering Baskets\nof All Sizes") + 
  theme(plot.title=element_text(size=10, face="bold", hjust = 0.5), 
        plot.subtitle=element_text(size=8, hjust = 0.5))

# removing the baskets above 30 items
plot2 <- private_item_cnt %>% 
  filter(private_item_cnt <= 30) %>%
  ggplot() + 
  geom_histogram(aes(private_item_cnt), binwidth=3) + 
  xlab("Private Label Items in Basket") + 
  ylab("Count") + 
  scale_y_continuous(labels=comma) + 
  ggtitle(label = "Distribution of\nPrivate Label Items\nper Basket" , 
          subtitle = "Excluding Baskets with\nGreater than 30 Items") + 
  theme(plot.title=element_text(size=10, face="bold", hjust = 0.5), 
        plot.subtitle=element_text(size=8, hjust = 0.5))

grid.arrange(plot1, plot2, ncol=2)
```

---

**Question 2**:  
The answer to Question 1 is all well and good, but people might only use the 
Brandefy app the first time they purchase a private label product. Repeat the 
analysis of Question 1 only counting the first-time purchase of a private label 
product per household. Hint: Arrange the transactions by `day`, then add a counter 
using the`row_number()` function to determine the first instance that a `product_id` 
was purchased by a `household_key`.

**Answer**:  
The average number of private label items per basket appears lower after deciding 
to only count the first time purchase of the item as the time when the app would 
be most likely used. If we used the data in Question 1 we might have been overstating 
the potential uses of the app.

```{r Q2, warning=FALSE, message=FALSE}
private_item_cnt_deduped <- transaction_data %>%
  inner_join(product, by='product_id') %>%
  arrange(day) %>%
  group_by(household_key, product_id) %>%
  mutate(purchase_instance = row_number()) %>%
  ungroup() %>%
  group_by(basket_id) %>%
  summarize(private_item_cnt = sum(brand == 'Private' & 
                                     purchase_instance == 1))

private_item_cnt_deduped %>% 
  filter(private_item_cnt <= 20) %>%
  ggplot() + 
  geom_histogram(aes(private_item_cnt), binwidth=1) + 
  xlab("First Purchase Private Label Items in Basket") + 
  ylab("Count") + 
  scale_y_continuous(labels=comma) + 
  ggtitle(label = "Distribution of Private Label Items per Basket" , 
          subtitle = "Only Considers First-Time Purchase of the Item (<= 20 items/basket)") + 
  theme(plot.title=element_text(size=10, face="bold"), 
        plot.subtitle=element_text(size=8))
```

---

**Question 3**:
Meg Greenhalgh saw the histogram you created for Question 2 that only considers 
first time purchases help and was impressed, but would also like some hard numbers. 
Calculate the following summary statistics: 

  1. The average number of private label products per basket
  2. The average sales value of these products per basket
  3. The percentage of baskets containing at least 1 product
  4. The percentage of baskets containing 0-2 products

**Answer**:
The answers to the questions posed are:

  1. 1.30 &mdash; the average number of private label products per basket
  2. $3.13 &mdash; the average sales value of these products per basket
  3. 45.6% &mdash; the percentage of baskets containing at least 1 product
  4. 83.8% &mdash; the percentage of baskets containing 0-2 products

```{r Q3, warning=FALSE, message=FALSE}
private_items_flagged <- transaction_data %>%
  inner_join(product, by='product_id') %>%
  arrange(day) %>%
  group_by(household_key, product_id) %>%
  mutate(purchase_instance = row_number()) %>%
  ungroup() %>%
  mutate(first_purchased_pvt_product = (brand == 'Private' & 
                                          purchase_instance == 1))

private_items_flagged %>%
  group_by(basket_id) %>%
  summarize(private_item_cnt = sum(first_purchased_pvt_product), 
            private_item_value = sum(sales_value[first_purchased_pvt_product])) %>%
  ungroup() %>%
  summarize(avg_private_items = mean(private_item_cnt), 
            avg_pvt_items_value = mean(private_item_value),
            perct_at_least_one = mean(private_item_cnt >= 1),
            perct_lte_two = mean(private_item_cnt <= 2))
```

---

**Question 4**:  
Meg Greenhalgh and her team have a limited amount of time and resources to review private 
label products, so they shifted towards a review club with category captains. 
What categories should Brandefy prioritize as it expands? To help answer this question, 
let's first try to prioritize based on the ratio of private to national label 
products. Meg should have good product review coverage in categories 
that have a high ratio of private label purchases over national brands. Create a 
list of the top 10 product categories based on ratio of private label to national 
product purchases. Here are some suggested steps to analyze:

 1. Inner Join `transaction_data` with `product` datasets to obtain the `commodity_desc` per transaction
 2. Obtain a count of transactions by `commodity_desc` and `brand` using the `count()` function
 3. Use the `spread()` function to create two columns of counts side-by-side (Private and National)
 4. Calculate a private-to-national ratio by dividing the `Private` column by the `National` column
 5. Calculate a column called `total_cnt` so we can filter out any noise 
 related to categories with only a handful of purchases.
 6. Filter the dataset to only the top 10 by private-to-national ratio categories 
 with at least 1,000 total transactions
 7. Arrange the data in descending order by the private-to-national ratio
 
**Answer**:  
Below is a list of the top 10 categories with frozen fruit, eggs, and milk topping 
the list. These products do not have strong brand loyalty, so shoppers are more 
likely to purchase the private label over the national brand.

```{r Q4}
transaction_data %>%
  inner_join(product, by='product_id') %>%
  count(commodity_desc, brand) %>%
  spread(brand, n) %>%
  mutate(total_cnt = Private + National, 
         private_ratio = Private / National) %>%
  filter(total_cnt > 1000, commodity_desc != 'COUPON/MISC ITEMS') %>%
  top_n(10, private_ratio) %>%
  arrange(desc(private_ratio))
```

---

**Question 5**:  
Meg Greenhalgh would also like to consider prioritizing product categories based on the price 
gap between national brands and private label brands. Create a list of the 
top 10 product categories based on price difference of the private and national brand 
product. Here are some suggested steps to analyze:

 1. Find the count of national and private label transactions in each `commodity_desc`
 2. Filter this table to only include categories with at least 100 national and 
 private label transactions. This can be used later to filter out and only look at 
 price differences between categories which have a critical mass of transactions in 
 both national and private label products.
 3. Inner Join `transaction_data` with `product` datasets to obtain the `commodity_desc` per transaction
 4. Inner Join the filtered counts data with the result from Step 3 by `commodity_desc` to 
 filter down the categories we're considering
 5. Group by `commodity_desc` and `brand` and calculate the average `sales value`
 6. Use the `spread()` function to put the private and national average prices side-by-side 
 and calculate the difference as `price_diff = Private - National`.
 7. Select the top 10 most negative prices using `top_n(-10, price_diff)`.

**Answer**: 
Infant formula, liquor, diapers, smoked meats, and laundry detergent top the list 
of products with the largest price difference between private label and national 
brands. It's not obvious to think of these product categories, but now that Meg Greenhalgh 
has this information, she can review the products and recommend them in her app 
as products having a really good value if the private label reviews well compared 
to the national label.

```{r Q5}
counts_data <- transaction_data %>%
  inner_join(product, by='product_id') %>%
  count(commodity_desc, brand) %>%
  spread(brand, n) %>%
  filter(Private > 100, National > 100)

transaction_data %>%
  inner_join(product, by='product_id') %>%
  inner_join(counts_data %>% 
               select(commodity_desc), by='commodity_desc') %>%
  group_by(commodity_desc, brand) %>%
  summarize(brand_cnt = length(brand), 
            avg_price = mean(sales_value, na.rm=TRUE)) %>%
  group_by(commodity_desc) %>%
  mutate(total_cnt = sum(brand_cnt)) %>%
  ungroup() %>%
  select(-brand_cnt) %>%
  spread(brand, avg_price) %>%
  mutate(price_diff = Private - National) %>%
  filter(is.finite(price_diff),
         commodity_desc != 'COUPON/MISC ITEMS') %>%
  top_n(-10, price_diff) %>%
  arrange(price_diff) %>%
  mutate(price_diff = dollar(price_diff))
```

---

**Question 6**:  
Meg Greenhalgh is interested in spending habits if the purchaser is also the 
user of the product. Meg has a theory that people purchase private 
label brands for their children but not for themselves in many occasions. The 
`completejourney` dataset provides a distinction for vitamins (children's vs. 
other types) that we can use to run a preliminary test on this hypothesis. Here 
are some suggested steps to analyze:

 1. Inner Join `transaction_data` with `product` datasets to obtain the `sub_commodity_desc` per transaction
 2. Filter to only the transactions with "VITAMIN" in the `sub_commodity_desc` field
 3. Summarize three metrics:  
   3.1 the total count of transactions being analyzed  
   3.2 the count that were private label transactions  
   3.3 the percentage of transactions that were private label  
 4. Filter out the noise by only considering sub-categories with over 100 total transactions

**Answer**:  
The data on vitamins supports that purchasers are less likely to purchase private label 
children's vitamins than any other types of vitamins. Only 16.1% of children's vitamins 
transactions were private label compared to 24.2% for all other vitamin types.

```{r, Q6a}
transaction_data %>%
  inner_join(product, by='product_id') %>%
  filter(grepl('VITAMIN', sub_commodity_desc)) %>%
  group_by(sub_commodity_desc) %>%
  summarize(total_cnt = length(brand), 
            private_cnt = sum(brand == 'Private'), 
            rate = mean(brand == 'Private')) %>%
  filter(total_cnt > 100) %>%
  arrange(rate)
```

```{r, Q6b}
transaction_data %>%
  inner_join(product, by='product_id') %>%
  filter(grepl('VITAMIN', sub_commodity_desc)) %>%
  mutate(childrens = ifelse(sub_commodity_desc == 'VITAMIN - CHILDREN', 'Yes', 'No')) %>%
  group_by(childrens) %>%
  summarize(total_cnt = length(brand), 
            private_cnt = sum(brand == 'Private'), 
            rate = mean(brand == 'Private')) %>%
  arrange(rate)
```

---

**Question 7**:  
In the previous question (Question 6) we determined that children's vitamins are 
the least likely to be purchased as private label compared to all other vitamin 
types. Run the same analysis again for products with `commodity_desc` = "BABYFOOD". 
Is this different from what you might have concluded in Question 6?

**Answer**:  
Baby food products have a 58.5% rate for purchasing private label. This contradicts 
our finding in the previous question where children's vitamins had a lower rate of 
private label buying that adult vitamins. We cannot make a blanket generalization 
that all products purchased for children are more likely to be private label. 
Brand loyalty is more likely the driving factor in each of the product segments.

```{r, Q7}
transaction_data %>%
  inner_join(product, by='product_id') %>%
  filter(commodity_desc == 'BABYFOOD') %>%
  group_by(commodity_desc) %>%
  summarize(total_cnt = length(brand), 
            private_cnt = sum(brand == 'Private'), 
            rate = mean(brand == 'Private')) %>%
  arrange(rate)
```
