---
title: "Data Transformation Exercises"
output: github_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The following five questions are based on concepts covered in 
[Chapter 5][chapter 5] of [R4DS][r4ds]. Answer them using the Complete Journey 
data and start by loading the `tidyverse` and `completejourney` packages.

```{r load-packages, warning=FALSE, message=FALSE}
library(tidyverse)
library(completejourney)
```

---

**Question 1**: Using `transaction_data`, create a two-column tibble sorted by 
basket, showing the most expensive item first. 

**Hint:** Sort by `basket_id`, then by `sales_value`. Return only those two
variables. 

This question is designed to strengthen your ability to use the `dplyr` 
verbs [`arrange()`][arrange] and [`select()`][select].

---

**Question 2**: Create two new variables named `regular_price` and 
`loyalty_price` according to the following logic:

-   `regular_price = (sales_value - (retail_disc + coupon_match_disc)) / quantity`
-   `loyalty_price = regular_price + (retail_disc / quantity)`

This question is designed to strengthen your ability to use the `dplyr` verb 
[`mutate()`][mutate].

---

**Question 3**: The `transaction_data` covers 92,339 unique product IDs. How 
many unique products (not transactions!) had a regular price of one dollar or 
less? What does this count equal for loyalty cardholders? 

**Hint:** After filtering select the product id column and count unique products 
using the `n_distinct()` function. 

This question is designed to strengthen your ability to use the `dplyr` verb 
[`filter()`][filter].

---

**Question 4**: What proportion of baskets are over $10 in sales value? 

**Hint:** You need to use `summarize()`, `group_by(basket)`, and `ungroup()`. 
Summarize over all baskets. In the last step you can calculate the proportion by 
taking the mean of `TRUE/FALSE` values. Use the code `mean(basket_value > 10)` 
to get the proportion over $10. 

This question is designed to strengthen your ability to use the `dplyr` verbs 
[`summarize()`][summarize] and [`ungroup()`][ungroup].

---

**Question 5**: Which store with over $10K in total `sales_value` discounts its 
products the most for loyal customers? 

**Hint:** You can calculate the loyalty discount as a percentage of regular 
price using the following logic:

-   `pct_loyalty_disc = 1 - (loyalty_price / regular_price)`

This question is designed to strengthen your ability to use the `dplyr` verbs 
[`filter()`][filter], [`mutate()`][mutate], [`group_by()`][summarize], 
[`summarize()`][summarize], and [`arrange()`][arrange] (i.e., 
almost everything covered in [Chapter 5][chapter5].

[chapter 5]: http://r4ds.had.co.nz/transform.html
[r4ds]: http://r4ds.had.co.nz/
[arrange]: http://r4ds.had.co.nz/transform.html#arrange-rows-with-arrange
[select]: http://r4ds.had.co.nz/transform.html#select-columns-with-select
[mutate]: http://r4ds.had.co.nz/transform.html#add-new-variables-with-mutate
[filter]: http://r4ds.had.co.nz/transform.html#filter-rows-with-filter
[summarize]: http://r4ds.had.co.nz/transform.html#grouped-summaries-with-summarise
[ungroup]: http://r4ds.had.co.nz/transform.html#ungrouping