---
title: "Data Visualization Exercises"
output: github_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The following five questions are based on concepts covered in [Chapter 3 of R4DS][Grolemund and Wickham 2017, 3]. They can be answered using the `transaction_data` and `product` datasets from the `completejourney` package. Start by loading the `tidyverse` and `completejourney` packages.

```{r load-packages, warning = FALSE, message = FALSE}
library(tidyverse)
library(completejourney)
```

---

**Question 1**: Create a histogram of `quantity`. What, if anything, do you find unusual about this visualization? 

This question is designed to strengthen your ability to use `geom_histogram()`.

---

**Question 2**: Use a line graph to plot total sales value by day. What, if anything, do you find unusual about this visualization?

This question is designed to strengthen your ability to use `dplyr` verbs in combination with `geom_line()`.

---

**Question 3**: Use a bar graph to compare the total sales values of national and private-label brands. 

**Hint**: Because `transaction_data` does not contain product metadata, run the code below to create a new dataset with additional product information in it. Use `my_transaction_data` for your answer.

```{r}
my_transaction_data <- left_join(transaction_data, product, by = 'product_id')
```

This question is designed to strengthen your ability to use `dplyr` verbs in combination with `geom_bar()` along with its `stat` argument.

---

**Question 4**: Building on Question 3, suppose you want to understand whether the retailer's customers' preference for national brands (compared to private-label brands) is stronger in the soft drink category than it is in the cheese category. Examine this supposition by using a stacked bar graph to compare the split between national and private-label brands for soft drinks and cheeses. 

**Hint**: Follow these three steps to create your plot: 

 - Filter `my_transaction_data` to include only transactions with `commodity_desc` equal to "SOFT DRINKS" or "CHEESE" 
 - Calculate total sales value by `commodity_desc` and `brand`
 - Create the bars using `geom_bar` with `stat = 'identity'` and `position = 'fill'`

---

**Question 5**: The code below filters `my_transaction_data` to include only peanut better, jelly, and jam transactions. Then it creates a new variable named `product_size` equal to product size in ounces. Create a bar graph with `pb_and_j_data` to visualize the distribution of the retailer's PB&J transactions by product size. Which two product sizes are the most popular?

```{r Q5, warning = FALSE}
pb_and_j_data <- my_transaction_data %>% 
  filter(commodity_desc == 'PNT BTR/JELLY/JAMS') %>%
  mutate(
    product_size = as.factor(as.integer(gsub('([0-9]+)([[:space:]]*OZ)', '\\1',
                                             curr_size_of_product)))
  )
```

[Grolemund and Wickham 2017, 3]: http://r4ds.had.co.nz/data-visualisation.html