---
title: "Exploratory Data Analysis (EDA) Exercises"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The following 7 questions are based on concepts covered in Chapters 7-8 in R4DS 
and can be answered using The Complete Journey data. Start by loading the `tidyverse` 
and the `completejourney` package.

```{r load-packages, warning=FALSE, message=FALSE}
library(tidyverse)
library(completejourney)
```

Run the following block of code to join datasets together and use when answering 
questions.

```{r data-prep}
my_transaction_data <- left_join(transaction_data, 
                                 product, by='product_id')
my_transaction_data <- left_join(my_transaction_data, 
                                 hh_demographic, by='household_key')
my_campaign_table <- left_join(campaign_table,
                               campaign_desc, by=c('campaign','description'))
```

---

**Question 1**: 
How many unique households exist in `my_transaction_data`? Also, how many of those 
households have demographic data in `hh_demographic`?

Here are some suggested steps: 

 1. Create a list of unique `household_key`s from `my_transaction_data` using `distinct()`
 2. Count of households using `nrow()`
 3. Use the `inner_join()` function to merge with `hh_demographic` 
 4. Count the remaining rows in your result
 
*This question grows your ability to count unique records in a dataset using `distinct()` and then use [`inner_join()`](http://r4ds.had.co.nz/relational-data.html#inner-join) to count records that existence in both datasets.*

---

**Question 2**: 
Determine the weekly average spend per person. Hint: Calculate total weekly sales 
value per household and divide by the number of people. Below is the code to convert 
`household_size_desc` into an integer. Remember, we only have household demographics 
for a select portion of the data, so your first step should be to filter to only 
transactions that have a non-missing value in `household_size_desc`. 

`mutate(hh_person_count = as.integer(gsub('\\+' ,'', household_size_desc)))`

*This question grows your ability to create a pipeline of data transformation steps to generate results. Specific functions include: [`filter()`](http://r4ds.had.co.nz/transform.html#filter-rows-with-filter), [`mutate()`](http://r4ds.had.co.nz/transform.html#add-new-variables-with-mutate), [`group_by()`, `summarize()`, and `ungroup()`](http://r4ds.had.co.nz/transform.html#grouped-summaries-with-summarise).*

---

**Question 3**:
Using the data derived in Question 2, plot the average spend per person segmented 
by count of children.  
*This question grows your ability to [`group_by()` and `summarize()`](http://r4ds.had.co.nz/transform.html#grouped-summaries-with-summarise) and then plot the result using `geom_bar()`.*

---

**Question 4**: 
Correlate spend on baby products with count of children. You can gauge the relationship 
visually by creating a scatterplot of the two variables and adding a trend line. 
Hint: This question requires transforming `kid_category_desc` similar to 
Question 2 when we transformed `household_size_desc` into an integer. 

Here are some suggested steps: 

  1. Filter out baby product transactions and non-missing household information on 
  children using this code: `filter(!is.na(kid_category_desc), grepl('BABY', sub_commodity_desc))`
  2. Create a variable of the number of children in a household using this code:  
  `mutate(hh_kid_count = as.integer(gsub('\\+' ,'', kid_category_desc)))`
  3. Summarize the total spend per child across households
  4. Create a scatterplot with line using `geom_point()` and `geom_smooth()`.
  
*This question grows your ability to [`filter()`](http://r4ds.had.co.nz/transform.html#filter-rows-with-filter), [`mutate()`](http://r4ds.had.co.nz/transform.html#add-new-variables-with-mutate), [`group_by()`, and `summarize()`](http://r4ds.had.co.nz/transform.html#grouped-summaries-with-summarise) and then plot the result using `geom_point()` and `geom_smooth()`.*

---

**Question 5**: 
In data analysis folklore it is said that Walmart increased beer sales by positioning 
diapers closer to the beer. What percentage of shopping baskets contained both 
diapers and beer?

---

**Question 6**:
Create a stacked bar chart showing across household income levels the percentage 
of sales from national brand products versus private label products. Describe the 
trend about the proportion of private label sales made across income levels.

---

**Question 7**: 
How often do retailers match the manufacturer's coupons? Hint: Filter the 
data where `coupon_disc > 0` and find the percentage of these observations that 
`coupon_match_disc < 0` is also true.
