---
title: "Data Transformation Solutions"
output: github_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The following five questions are based on concepts covered in 
[Chapter 5][chapter 5] of [R4DS][r4ds]. Answer them using `transaction_data` in
the Complete Journey data package and start by loading the `tidyverse` and 
`completejourney` packages.

```{r load-packages, warning = FALSE, message = FALSE}
library(tidyverse)
library(completejourney)

transaction_data <- transaction_data %>% 
  select(
    quantity,
    sales_value, 
    retail_disc, coupon_disc, coupon_match_disc,
    household_key, store_id, basket_id, product_id, 
    week_no, day, trans_time
  )
```

---

**Question 1**: Change the discount variables (i.e., `retail_disc`, 
`coupon_disc`, `coupon_match_disc`) from negative to positive. 

**Hint:** Use the `abs()` function within [`mutate()`][mutate]. 

This question is designed to strengthen your ability to use the `dplyr` 
verb [`mutate()`][mutate] to overwrite an existing variable.

**Answer**:
```{r}
transaction_data <- transaction_data %>% 
  mutate(
    retail_disc       = abs(retail_disc),
    coupon_disc       = abs(coupon_disc),
    coupon_match_disc = abs(coupon_match_disc)
  )
```

---

**Question 2**: Create three new variables named `regular_price`, 
`loyalty_price`, and `coupon_price` according to the following logic:

```{r eval = FALSE}
regular_price = (sales_value + retail_disc + coupon_match_disc) / quantity
loyalty_price = (sales_value + coupon_match_disc) / quantity
coupon_price  = (sales_value - coupon_disc) / quantity
```


This question is designed to strengthen your ability to use the `dplyr` verb 
[`mutate()`][mutate] to create new variables from existing ones. It should also
help you develop a better understanding of the discount variables in 
`transaction_data`.

**Answer:**
```{r}
(transaction_data <- transaction_data %>% 
  mutate(
    regular_price     = (sales_value + retail_disc + coupon_match_disc) / 
                        quantity,
    loyalty_price     = (sales_value + coupon_match_disc) / 
                        quantity,
    coupon_price      = (sales_value - coupon_disc) / 
                        quantity
  ) %>% 
  select(regular_price, loyalty_price, coupon_price, everything())
)
```

---

**Question 3**: `transaction_data` includes 92,339 unique product IDs. How many 
of these products (not transactions!) had a regular price of one dollar or less? 
What does this count equal for loyalty and coupon prices?

**Hint:** After filtering, select the `product_id` column and count unique 
products using the `n_distinct()` function. 

This question is designed to strengthen your ability to use the `dplyr` verbs 
[`filter()`][filter] and [`select()`][select].

**Answer:**

12,442 products had a regular price less than or equal to $1.00. This count for 
loyalty price is 20,113 products. For coupon price, it's 22,273. These numbers 
indicate that over 7,500 products were discounted to a dollar or less with 
loyalty-card promotions or coupons.

```{r}
transaction_data %>% 
  filter(regular_price <= 1) %>% 
  select(product_id) %>% 
  n_distinct()

transaction_data %>% 
  filter(loyalty_price <= 1) %>% 
  select(product_id) %>% 
  n_distinct()

transaction_data %>% 
  filter(coupon_price <= 1) %>% 
  select(product_id) %>% 
  n_distinct()
```

---

**Question 4**: What proportion of baskets are over $10 in sales value? 

**Hint:** You need to use [`group_by()`][summarize], [`summarize()`][summarize], 
and [`ungroup()`][ungroup]. Summarize over all baskets. In the last step, you 
can calculate the proportion by taking the mean of `TRUE/FALSE` values. Use 
`mean(basket_value > 10)` to get the proportion over $10. 

This question is designed to strengthen your ability to use the `dplyr` verbs 
[`group_by()`][summarize], [`summarize()`][summarize] and [`ungroup()`][ungroup].

**Answer:**

Approximately 65% of baskets bring the retailer over $10 in value.

```{r}
transaction_data %>%
  group_by(basket_id) %>%
  summarize(basket_value = sum(sales_value)) %>%
  ungroup() %>%
  summarize(proportion_over_10 = mean(basket_value > 10))
```

---

**Question 5**: Which store with over $10K in total `sales_value` discounts its 
products the most for loyal customers? 

**Hint:** You can calculate loyalty discount as a percentage of regular price 
using the following logic:

```{r eval = FALSE}
pct_loyalty_disc = 1 - (loyalty_price / regular_price)
```

This question is designed to strengthen your ability to use the `dplyr` verbs 
[`filter()`][filter], [`mutate()`][mutate], [`group_by()`][summarize], 
[`summarize()`][summarize], and [`arrange()`][arrange] (i.e., 
almost everything covered in [Chapter 5][chapter 5].

**Answer:**

`store_id == 341` has, on average, a discount of 18.8% off regular price for 
members of the retailer's loyalty program. The code below returns all stores in 
rank order to give a sense of how similar the discounting is (not) between 
stores.
```{r}
transaction_data %>%
  filter(
    is.finite(regular_price), 
    is.finite(loyalty_price), 
    regular_price > 0
  ) %>%
  mutate(
    pct_loyalty_disc     = 1 - (loyalty_price / regular_price)
  ) %>%
  group_by(store_id) %>%
  summarize(
    total_sales_value    = sum(sales_value), 
    avg_pct_loyalty_disc = mean(pct_loyalty_disc)
  ) %>%
  filter(total_sales_value > 10000) %>%
  arrange(desc(avg_pct_loyalty_disc))
```

[chapter 5]: http://r4ds.had.co.nz/transform.html
[r4ds]: http://r4ds.had.co.nz/
[arrange]: http://r4ds.had.co.nz/transform.html#arrange-rows-with-arrange
[select]: http://r4ds.had.co.nz/transform.html#select-columns-with-select
[mutate]: http://r4ds.had.co.nz/transform.html#add-new-variables-with-mutate
[filter]: http://r4ds.had.co.nz/transform.html#filter-rows-with-filter
[summarize]: http://r4ds.had.co.nz/transform.html#grouped-summaries-with-summarise
[ungroup]: http://r4ds.had.co.nz/transform.html#ungrouping