---
title: "Data Wrangling Exercises"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The following four exercises are based on concepts covered in 
[Chapter 12][G&W 2017, 12] and [Chapter 13][G&W 2017, 13] of 
[R for Data Science][G&W 2017]. Use the `coupon`, `coupon_redempt`, 
`campaign_table`, `transaction_data`, and `product` datasets that come with the 
`completejourney` package to work on them, and start by loading the `tidyverse` 
and `completejourney` packages.

```{r, echo = TRUE, eval = TRUE, message = FALSE}
library(tidyverse)
library(completejourney)
```

---

## Exercise 1
What percent of households that received the retailer's weekly mailer redeemed 
at least one coupon?


### Solution to Exercise 1
Of the 1,584 households that were mailed coupons as part of a campaign, 434 (or
27.4%) redeemed a coupon.

```{r}
left_join(
  campaign_table %>% count(household_key),
  coupon_redempt %>% count(household_key), by = 'household_key'
) %>% 
  summarize(redemption_rate = mean(!is.na(n.y)))
```

---

## Exercise 2
How many households did not redeem a coupon?   

### Solution to Exercise 2
1,150 households did not redeem a coupon. 

```{r}
left_join(
  campaign_table %>% count(household_key),
  coupon_redempt %>% count(household_key), by = 'household_key'
) %>% 
  summarize(redemption_rate = sum(is.na(n.y)))
```

---

## Exercise 3
What percent of coupons promoted in the retailer's weekly mailer got redeemed at
least once?

### Solution to Exercise 3
Of the 1,135 coupons promoted in the retailer's weekly mailer, 556 (or 49%)
coupons were redeemed at least once.

```{r}
left_join(
  coupon %>% count(coupon_upc),
  coupon_redempt %>% count(coupon_upc), by = 'coupon_upc'
) %>% 
  summarize(redemption_rate = mean(!is.na(n.y)))
```

---

## Exercise 4
Using the `transaction_data` and `product` datasets, determine which product
category (i.e., `sub_commodity_desc`) grew the most in terms of revenue for the
retailer in the second half of the study period (i.e., between `week_no == 52` 
and `week_no == 102`). Only consider product categories that had over $100 in
revenue in week 52, and calculate revenue growth as a percentage of category
revenue in week 52.

Here are some suggested steps: 

 1. Join the `transaction_data` and `product` datasets.
 2. Group the data by `sub_commodity_desc` and `week_no`.
 3. Calculate `sales_value` at the category level.
 4. Filter the data to only include category revenues in weeks 52 and 102.
 5. Create a new variable called `revenue_growth`, making use of the `lag()`
 function. Recognize that you only want to consider categories that had category
 revenues of at least $100 in week 52.
 6. Arrange the data in descending order according to `revenue_growth`.

### Solution to Exercise 4
Having grown 85.2% in the second half of the study period, the **select beef**
category grew the most among all product categories.

```{r, message = FALSE}
left_join(transaction_data, product) %>%
  group_by(sub_commodity_desc, week_no) %>%
  summarize(sales_value_category = sum(sales_value, na.rm = TRUE)) %>%
  filter(week_no == 52 | week_no == 102) %>%
  mutate(
    revenue_growth = ifelse(lag(sales_value_category) >= 100,
                            (sales_value_category - lag(sales_value_category)) /
                             lag(sales_value_category), NA)
  ) %>%
  arrange(desc(revenue_growth))
```

[G&W 2017]: http://r4ds.had.co.nz/
[G&W 2017, 12]: http://r4ds.had.co.nz/tidy-data.html
[G&W 2017, 13]: http://r4ds.had.co.nz/relational-data.html