---
title: "Brandefy Private Label Case Solutions"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The following 6 questions are based on the Brandefy case study and can be answered 
using The Complete Journey data. Start by loading the `tidyverse` and the 
`completejourney` package.

```{r load-packages, warning=FALSE, message=FALSE}
# required packages
library(tidyverse)
library(completejourney)
```

---

**Question 1**:  
Meg is trying to gauge the overall opportunity for usage of an app that shoppers 
could use in-store to scan items and see product reviews of similar private label 
products. To get a sense for potential app usage count the number of private 
label items per basket and plot the distribution as a histogram. Hint: When calculating 
the average private label items per basket ensure you include the baskets with zero 
private label items. You can do this by summarizing the baskets using:  
`summarize(private_item_cnt = sum(brand == 'Private'))`.

**Answer**:  
The majority of baskets have between 0-2 private label items. The average is 2.7 
items; however, that is because it is a very long-tailed distribution. This is an 
interesting insight because Meg may be able to sub-segment potential users of the 
app into casual buyers and die-hard private label fans.

```{r Q1, warning=FALSE, message=FALSE}
private_item_cnt <- transaction_data %>%
  inner_join(product, by='product_id') %>%
  group_by(basket_id) %>%
  summarize(private_item_cnt = sum(brand == 'Private'))

private_item_cnt %>% 
  summarize(avg_private_item_cnt = mean(private_item_cnt))

ggplot(private_item_cnt) + 
  geom_histogram(aes(private_item_cnt))

# removing the baskets above 30 items
private_item_cnt %>% 
  filter(private_item_cnt <= 30) %>%
  ggplot() + 
  geom_histogram(aes(private_item_cnt), binwidth=1)
```

---

**Question 2**:  
Since shoppers might only use the app for a new purchase (rather than a repeat buy) 
repeat the analysis of Question 1 and only count first-time purchases of private label 
products per household. Hint: Arrange the transactions by `day`, then run 
`distinct(household_key, product_id, .keep_all=TRUE)` to get the first time that 
a household purchased a product.

**Answer**:  
The average number of private label items per basket drops to 1.65 when only considering 
first time product purchases.

```{r Q2, warning=FALSE, message=FALSE}
private_item_cnt_deduped <- transaction_data %>%
  inner_join(product, by='product_id') %>%
  arrange(day) %>%
  distinct(household_key, product_id, .keep_all=TRUE) %>%
  group_by(basket_id) %>%
  summarize(private_item_cnt = sum(brand == 'Private'))

private_item_cnt_deduped %>% 
  summarize(avg_private_item_cnt = mean(private_item_cnt))

# removing the baskets above 20 items
private_item_cnt_deduped %>% 
  filter(private_item_cnt <= 20) %>%
  ggplot() + 
  geom_histogram(aes(private_item_cnt), binwidth=1)
```

---

**Question 3**:  
Meg and her team have a limited amount of time and resources to review private 
label products. What categories should Brandefy prioritize as it expands? To answer 
this question, let's first try to prioritize based on the ratio of private to national 
label products. Meg should have good product review coverage in categories that have a 
high ratio of private label purchases over national brands. Create a list of the 
top 10 product categories based on ratio of private label to national product purchases. 
Here are some suggested steps to analyze:

 1. Inner Join `transaction_data` with `product` datasets to obtain the `commodity_desc` per transaction
 2. Obtain a count of transactions by `commodity_desc` and `brand` using the `count()` function
 3. Use the `spread()` function to create two columns of counts side-by-side (Private and National)
 4. Calculate a private-to-national ratio by dividing the `Private` column by the `National` column
 5. Calculate a column called `total_cnt` so we can filter out any noise 
 related to categories with only a handful of purchases.
 6. Filter the dataset to only the top 10 by private-to-national ratio categories 
 with at least 1,000 total transactions
 7. Arrange the data in descending order by the private-to-national ratio
 
**Answer**:  
Below is a list of the top 10 categories with frozen fruit, eggs, and milk topping 
the list. These products do not have strong brand loyalty, so shoppers are more 
likely to purchase the private label over the national brand.

```{r Q3}
transaction_data %>%
  inner_join(product, by='product_id') %>%
  count(commodity_desc, brand) %>%
  spread(brand, n) %>%
  mutate(total_cnt = Private + National, 
         private_ratio = Private / National) %>%
  filter(total_cnt > 1000, commodity_desc != 'COUPON/MISC ITEMS') %>%
  top_n(10, private_ratio) %>%
  arrange(desc(private_ratio))
```

---

**Question 4**:  
Meg would also like to consider prioritizing product categories based on the price 
gap between national brands and private label brands. Create a list of the 
top 10 product categories based on price difference of the private and national brand 
product. Here are some suggested steps to analyze:

 1. Find the count of national and private label transactions in each `commodity_desc`
 2. Filter this table to only include categories with at least 100 national and 
 private label transactions. This can be used later to filter out and only look at 
 price differences between categories which have a critical mass of transactions in 
 both national and private label products.
 3. Inner Join `transaction_data` with `product` datasets to obtain the `commodity_desc` per transaction
 4. Inner Join the filtered counts data with the result from Step 3 by `commodity_desc` to 
 filter down the categories we're considering
 5. Group by `commodity_desc` and `brand` and calculate the average `sales value`
 6. Use the `spread()` function to put the private and national average prices side-by-side 
 and calculate the difference as `price_diff = Private - National`.
 7. Select the top 10 most negative prices using `top_n(-10, price_diff)`.

**Answer**: 
Infant formula, liquor, diapers, smoked meats, and laundry detergent top the list 
of products with the largest price difference between private label and national 
brands. It's not obvious to think of these product categories, but now that Meg 
has this information, she can review the products and recommend them in her app 
as products having a really good value if the private label reviews well compared 
to the national label.

```{r Q4}
counts_data <- transaction_data %>%
  inner_join(product, by='product_id') %>%
  count(commodity_desc, brand) %>%
  spread(brand, n) %>%
  filter(Private > 100, National > 100)

transaction_data %>%
  inner_join(product, by='product_id') %>%
  inner_join(counts_data %>% 
               select(commodity_desc), by='commodity_desc') %>%
  group_by(commodity_desc, brand) %>%
  summarize(brand_cnt = length(brand), 
            avg_price = mean(sales_value, na.rm=TRUE)) %>%
  group_by(commodity_desc) %>%
  mutate(total_cnt = sum(brand_cnt)) %>%
  ungroup() %>%
  select(-brand_cnt) %>%
  spread(brand, avg_price) %>%
  mutate(price_diff = Private - National) %>%
  filter(is.finite(price_diff),
         commodity_desc != 'COUPON/MISC ITEMS') %>%
  top_n(-10, price_diff) %>%
  arrange(price_diff)
```

---

**Question 5**:  
Meg is interested in spending habits if the purchaser is also the user of the product. 
Meg has a theory that people purchase private label brands for their children but not for 
themselves in many occasions. The `completejourney` dataset provides a distinction 
for vitamins (children's vs. other types) that we can use to run a preliminary test 
on this hypothesis. Here are some suggested steps to analyze:

 1. Inner Join `transaction_data` with `product` datasets to obtain the `sub_commodity_desc` per transaction
 2. Filter to only the transactions with "VITAMIN" in the `sub_commodity_desc` field
 3. Summarize three metrics:  
   3.1 the total count of transactions being analyzed  
   3.2 the count that were private label transactions  
   3.3 the percentage of transactions that were private label  
 4. Filter out the noise by only considering sub-categories with over 100 total transactions

**Answer**:  
The data on vitamins supports that purchasers are less likely to purchase private label 
children's vitamins than any other types of vitamins. Only 16.1% of children's vitamins 
transactions were private label compared to 24.2% for all other vitamin types.

```{r, Q5a}
transaction_data %>%
  inner_join(product, by='product_id') %>%
  filter(grepl('VITAMIN', sub_commodity_desc)) %>%
  group_by(sub_commodity_desc) %>%
  summarize(total_cnt = length(brand), 
            private_cnt = sum(brand == 'Private'), 
            rate = mean(brand == 'Private')) %>%
  filter(total_cnt > 100) %>%
  arrange(rate)
```

```{r, Q5b}
transaction_data %>%
  inner_join(product, by='product_id') %>%
  filter(grepl('VITAMIN', sub_commodity_desc)) %>%
  mutate(childrens = ifelse(sub_commodity_desc == 'VITAMIN - CHILDREN', 'Yes', 'No')) %>%
  group_by(childrens) %>%
  summarize(total_cnt = length(brand), 
            private_cnt = sum(brand == 'Private'), 
            rate = mean(brand == 'Private')) %>%
  arrange(rate)
```

---

**Question 6**:  
In the previous question we determined that children's vitamins are the least likely 
to be purchased as private label compared to all other vitamin types. Run the same 
analysis again for products with `commodity_desc` = "BABYFOOD". Is this different 
from what you might have concluded in Question 5?

**Answer**:  
Baby food products have a 58.5% rate for purchasing private label. This contradicts the 
findings in the previous question where vitamins had a low rate of private label buying. 
We cannot make a blanket generalization that all products purchased for children are 
more likely to be private label. Brand loyalty is more likely the driving factor in 
each of the product segments.

```{r, Q6}
transaction_data %>%
  inner_join(product, by='product_id') %>%
  filter(commodity_desc == 'BABYFOOD') %>%
  group_by(commodity_desc) %>%
  summarize(total_cnt = length(brand), 
            private_cnt = sum(brand == 'Private'), 
            rate = mean(brand == 'Private')) %>%
  arrange(rate)
```
